### 2009-04-22 docs/flowchart/flowchart.tex
more work on spherical base state

\item {\bf Advect Base Pressure}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow$ 
$[p_0^\outp,\psi^\outp]$\\
is the process by which we update the base state pressure through $\dt$ in time and
compute $\psi$.
\begin{description}

\item[plane-parallel:] First we compute a cell-centered $\psi^\outp$ using
\begin{equation}
 \psi_j^\outp = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
\end{equation}
Then, we discretize the base state pressure evolution equation (paper III; eq.~[31]),
\begin{equation}
\frac{\partial p_0}{\partial t} = 
-w_0 \frac{\partial p_0}{\partial r} + \psi \enskip ,
\label{eq:p0upd_new}
\end{equation}
to compute the new base state pressure,
\begin{equation}
p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\outp 
\enskip,
\end{equation}
where the interface states are again found via the procedure described
in paper II, Appendix A.

\item[spherical:] See Section \ref{Sec:Advect Base Spherical} for working notes.

@@

\item[spherical:] 
\MarginPar{Need to write this up somewhere}

>>>>

\item {\bf Advect Base Pressure}
is the process by which we update the base state pressure through $\dt$ in time and
compute $\psi$.  The input arguments are different for plane-parallel and spherical.
\begin{description}
\item {\bf Advect Base Pressure Planar}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow 
[p_0^\outp,\psi^\outp]$:

First we compute a cell-centered $\psi^\outp$ using
\begin{equation}
 \psi_j^\outp = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
\end{equation}
Then, we discretize the base state pressure evolution equation (paper III; eq.~[31]),
\begin{equation}
\frac{\partial p_0}{\partial t} = 
-w_0 \frac{\partial p_0}{\partial r} + \psi \enskip ,
\label{eq:p0upd_new}
\end{equation}
to compute the new base state pressure,
\begin{equation}
p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\outp 
\enskip,
\end{equation}
where the interface states are again found via the procedure described
in paper II, Appendix A.\\

\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},\Sbar^{\inp}] 
\rightarrow [p_0^{\outp},\psi^{\outp}]$:

First use a predictor-corrector formulation for the pressure update 
in order to improve the overall accuracy.  Beginning with equation~(\ref{eq:w01}),
we can write
\begin{eqnarray*}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
 \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
\end{eqnarray*}
We time-center the left hand side:
\[
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
\]
We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
\begin{equation}
p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
\end{equation}
with:
\begin{equation}
f = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
\end{equation}
where the $(\partial p_0^\inp / \partial r)$ is computed by
interpolating $p_0^\inp$ from cell-centers to edges and then
differencing.  Now we begin the corrector step.  Using $p_0^\star$, we can compute a more 
accurate approximation to equation~(\ref{eq:frhs}),
\begin{equation}
\tilde{f} = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} \tilde{p_0}} w_0^{\inp} \left ( \frac{\partial \tilde{p_0}}{\partial r} \right ) \right ] \enskip ,
\end{equation}
where $\tilde{p} = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
by first putting $\tilde{p}$ on edges.

This allows us to complete the corrector for the pressure update:
\begin{equation}
p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} \tilde{f}\right)}{\left(1 - \gammabar^{\inp} \tilde{f}\right)}\enskip .
\end{equation}
We note that since we do not time-center $\gammabar$, this is not
strictly second-order accurate.  However, as $\gammabar$ varies slowly
in time, we expect the effect to be small.  This is something that we 
defer until simulations indicate it is warranted.

Next, we compute $\psi^{\outp}$.  Rewriting Equation (\ref{eq:w01}), we can define
\begin{eqnarray}
\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
\end{eqnarray}
We discretize this locally as:
\begin{equation}
\psi_j^{\outp} = 
\left[\gammabar^{\inp} \frac{\left( p_0^{\inp} + p_0^{\outp} \right)}{2}\right]_j
\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
          (r^2 w_0^{\outp})_{j-\myhalf} \right ] \right \} \enskip .
\end{equation}
\end{description}

@@

### 2009-04-22 docs/spherical_basestate/basestate.tex
more work on spherical base state

We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
corresponds to $i=1,$ by setting:
\begin{eqnarray}
A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
B_1 &=& 1\enskip. \nonumber
\end{eqnarray}
We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $i=N,$
by setting:
\begin{eqnarray}
A_N &=& -1\enskip, \nonumber \\
B_N &=&  1\enskip, \nonumber \\
C_N = F_N &=& 0\enskip. \nonumber
\end{eqnarray}
\subsection{Calculation of $\psi$}

Rewriting Equation (\ref{eq:w01}), we can define
\begin{eqnarray}
\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
\end{eqnarray}
We discretize this locally as:
\begin{eqnarray}
\psi_j^{\nph,\star} &=& \gammabar^{(1)} \frac{\left( p_0^{n} + p_0^{n+1,\star} \right)_j}{2}
\left \{ \overline{S^{\nph,\star}_j} - 
 \frac{1}{r_j^2} \left [ (r^2 {w_0}^{\nph,\star})_{j+\myhalf} -
          (r^2 {w_0}^{\nph,\star})_{j-\myhalf} \right ] \right \} \enskip , \nonumber \\
&& \\
\psi_j^{\nph} &=& \gammabar^{n+1,\star} \frac{\left( p_0^{n} + p_0^{n+1} \right)_j}{2}
\left \{ \overline{S^{\nph}_j} - 
 \frac{1}{r_j^2} \left [ (r^2 {w_0}^{\nph})_{j+\myhalf} -
          (r^2 {w_0}^{\nph})_{j-\myhalf} \right ] \right \} \enskip .
\end{eqnarray}
\MarginPar{Why is $\gammabar$ not time centered in the $\psi_j^{\nph}$ equation?}

\subsection{{\bf Advect Base}}\label{Sec:Advect Base Spherical}

\begin{enumerate}

\item {\bf Pressure Update:}
\newline
We use a predictor-corrector formulation for the pressure update 
in order to improve the overall accuracy.  Beginning with equation~(\ref{eq:w01}),
we can write
\begin{eqnarray*}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
 \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
\end{eqnarray*}
We time-center the left hand side:
\[
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
\]
We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
\begin{equation}
p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
\end{equation}
with:
\begin{equation}
f = \frac{\dt}{2} \left [ \overline{S^{\inp}} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
\end{equation}
where the $(\partial p_0^\inp / \partial r)$ is computed by
interpolating $p_0^\inp$ from cell-centers to edges and then
differencing.

Now we begin the corrector step.  Using $p_0^\star$, we can compute a more accurate approximation
to equation~(\ref{eq:frhs}),
\begin{equation}
\tilde{f} = \frac{\dt}{2} \left [ \overline{S^{\inp}} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} \tilde{p_0}} w_0^{\inp} \left ( \frac{\partial \tilde{p_0}}{\partial r} \right ) \right ] \enskip ,
\end{equation}
where $\tilde{p} = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
by first putting $\tilde{p}$ on edges.

This allows us to complete the corrector for the pressure update:
\begin{equation}
p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} \tilde{f}\right)}{\left(1 - \gammabar^{\inp} \tilde{f}\right)}\enskip .
\end{equation}
We note that since we do not time-center $\gammabar$, this is not
strictly second-order accurate.  However, as $\gammabar$ varies slowly
in time, we expect the effect to be small.  This is something that we 
defer until simulations indicate it is warranted. \MarginPar{new}

\end{enumerate}

\subsection{\bf{Correct Base}}

The is the process by which we adjust the base state density to account for
large-scale mixing.  It can also be used to adjust the base state enthalpy
to account for large-scale mixing, reactions, and thermal diffusion.
We have simplified the algorithm greatly by simply setting $\nabla\cdot\etarho$
equal to the averaged difference between the base state and full state.

\begin{equation}
\rho_0^{\outp} = \rho_0^{\inp} + \underbrace{{\rm{\bf Avg}}(\rho^{\inp} - \rho_0^{\inp})}_{-\Delta t\nabla\cdot\etarho}.
\end{equation}

>>>>

We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
corresponds to $i=1,$ by setting:
\begin{eqnarray}
A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
B_1 &=& 1\enskip. \nonumber
\end{eqnarray}
We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $i=N,$
by setting:
\begin{eqnarray}
A_N &=& -1\enskip, \nonumber \\
B_N &=&  1\enskip, \nonumber \\
C_N = F_N &=& 0\enskip. \nonumber
\end{eqnarray}
